---
title: "HW3"
author: Zhihui Zhang
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(class)
library(ISLR2)
data("Weekly")
library(MASS)
library(e1071)
library(dplyr)
library(leaps)
library(pls)
options(scipen = 999)
```



## 13. This question should be answered using the Weekly data set, which is part of the ISLR2 package. This data is similar in nature to the Smarket data from this chapter’s lab, except that it contains 1, 089 weekly returns for 21 years, from the beginning of 1990 to the end of 2010.
## (a) Produce some numerical and graphical summaries of the Weekly data. Do there appear to be any patterns?

```{r}
summary(Weekly)
pairs(Weekly)
cor(Weekly[, -9])
```
Yes, the are some patterns. Like in Lag2, the direction have different group mean in lag2 predictors. The similar pattern also shows in Lag3 and Lag4.

## (b) Use the full data set to perform a logistic regression with Direction as the response and the five lag variables plus Volume as predictors. Use the summary function to print the results. Do any of the predictors appear to be statistically significant? If so, which ones?

```{r}
mod4.13 <- glm(Direction ~.-Year - Today, family = binomial(), data = Weekly )
summary(mod4.13)
```
The Lag2 os statistically significant. 

## (c) Compute the confusion matrix and overall fraction of correct predictions. Explain what the confusion matrix is telling you about the types of mistakes made by logistic regression.

```{r}
mod4.13.probs <- predict(mod4.13, type = "response")
mod4.13.pred <- rep("Down", length(Weekly$Direction))
mod4.13.pred[mod4.13.probs > .5] = "Up"
table(Weekly$Direction, mod4.13.pred)
```

The cell in the up-right give us the false positive case which their true value are negative while we predict them to be positive. The left-down give us the false negative which we falsely predict the positive case to be negative. 


## (d) Now fit the logistic regression model using a training data period from 1990 to 2008, with Lag2 as the only predictor. Compute the confusion matrix and the overall fraction of correct predictions for the held out data (that is, the data from 2009 and 2010).

```{r}
train_set <- Weekly %>% filter(Year >= 1990 & Year <= 2008)
test_set <- Weekly %>% filter(Year %in% c(2009, 2010))
mod4.13d <- glm(Direction ~ Lag2, family = binomial(), data = train_set)
summary(mod4.13d)

mod4.13d.probs <- predict(mod4.13d, test_set, type = 'response')
mod4.13d.pred <- ifelse(mod4.13d.probs > 0.5, 'Up', 'Down')
#confusion matrix for the test data
table(test_set$Direction, mod4.13d.pred)
#correct prediction
mean(mod4.13d.pred == test_set$Direction)
```

## (e) Repeat (d) using LDA.

```{r}
mod4.13e <- lda(Direction ~ Lag2, data = train_set)
mod4.13e
mod4.13e.pred <- predict(mod4.13e, test_set)
#names(mod4.13e.pred)
mod4.13e.class <- mod4.13e.pred$class
table(mod4.13e.class, test_set$Direction)
#correct prediction
mean(mod4.13e.class == test_set$Direction)
plot(mod4.13e)
```

## (f) Repeat (d) using QDA.

```{r}
mod4.13f <- qda(Direction ~ Lag2, data = train_set)
mod4.13f
mod4.13f.pred <- predict(mod4.13f, test_set)
#names(mod4.13f.pred)
mod4.13f.class <- mod4.13f.pred$class
table(mod4.13f.class, test_set$Direction)
#correct prediction
mean(mod4.13f.class == test_set$Direction)
```
## (g) Repeat (d) using KNN with K = 1.

```{r}

train.X <- data.frame(train_set$Lag2)
test.X <-  data.frame(test_set$Lag2)
train.Direction <- train_set$Direction
test.Direction <- test_set$Direction
set.seed(1)
mod4.13g <- knn(train.X, test.X, train.Direction, k = 1)
table(mod4.13g, test.Direction)
mean(mod4.13g == test.Direction)
```

## (h) Repeat (d) using naive Bayes.

```{r}
mod4.13h <- naiveBayes(Direction ~ Lag2, data = train_set)
mod4.13h
mod4.13h.class <- predict(mod4.13h, test_set)
table(mod4.13h.class, test_set$Direction)
mean(mod4.13h.class == test_set$Direction)
```

## (i) Which of these methods appears to provide the best results on this data?

The logistic regression and LDA provide the best results

## (j) Experiment with different combinations of predictors, including possible transformations and interactions, for each of the methods. Report the variables, method, and associated confusion matrix that appears to provide the best results on the held out data. Note that you should also experiment with values for K in the KNN classifier.

```{r}
#logistic regression
log.mod <- glm(Direction ~ Lag2 + Lag3 + Lag4 + Lag5, family = binomial(), data = train_set)
summary(log.mod)
log.prob <- predict(log.mod, test_set, type = 'response')
log.pred <- ifelse(log.prob > 0.5, 'Up', 'Down')
table(log.pred, test_set$Direction)
mean(log.pred == test_set$Direction)
```

```{r}
#LDA
lda.mod <- lda(Direction ~ Lag2 + Lag3 + Lag4 + Lag5, data = train_set)
lda.pred <- predict(lda.mod, test_set)
lda.class <- lda.pred$class
table(lda.class, test_set$Direction)
mean(lda.class == test_set$Direction)
```

```{r}
#QDA
qda.mod <- qda(Direction ~ Lag2 + Lag3, data = train_set)
qda.pred <- predict(qda.mod, test_set)
qda.class <- qda.pred$class
table(qda.class, test_set$Direction)
mean(qda.class == test_set$Direction)
```

```{r}
#KNN
train.X <- data.frame(train_set$Lag2, train_set$Lag3)
test.X <-  data.frame(test_set$Lag2, test_set$Lag3)
train.Direction <- train_set$Direction
test.Direction <- test_set$Direction
knn.mod <- knn(train.X, test.X, train.Direction, k = 6 )
table(knn.mod, test.Direction)
mean(knn.mod == test.Direction)
```

```{r}
nb.mod <- naiveBayes(Direction ~ Lag2 + Lag3 , data = train_set)
nb.mod.class <- predict(nb.mod, test_set)
table(nb.mod.class, test_set$Direction)
mean(nb.mod.class == test_set$Direction)
```

The interaction term will not work, since the correlation among two predictors are extremley low. And the transformation of predictors also will not help a lot to improve the predictions. 



## 14. In this problem, you will develop a model to predict whether a given car gets high or low gas mileage based on the Auto data set.

## (a) Create a binary variable, mpg01, that contains a 1 if mpg contains a value above its median, and a 0 if mpg contains a value below its median. You can compute the median using the median() function. Note you may find it helpful to use the data.frame() function to create a single data set containing both mpg01 and the other Auto variables.

```{r}
auto <- Auto
auto$mpg01 <- ifelse(Auto$mpg > median(Auto$mpg), 1, 0)
```

## (b) Explore the data graphically in order to investigate the association between mpg01 and the other features. Which of the other features seem most likely to be useful in predicting mpg01? Scatterplots and boxplots may be useful tools to answer this question. Describe your findings.

```{r}
pairs(auto[-1])
boxplot(auto$mpg01, auto$displacement)
boxplot(auto$mpg01, auto$horsepower)
cor(auto[-c(1,9)])
```

From the scatter plots and box plots above, we could see that displacement, horsepower, weight and acceleration might relate to mpg01. 

## (c) Split the data into a training set and a test set.

```{r}
set.seed(123)
smp_size <- floor(0.70 * nrow(auto))
train_ind <- sample(seq_len(nrow(auto)), size = smp_size)

train <- auto[train_ind, ]
test <- auto[-train_ind, ]
```

## (d) Perform LDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?

```{r}
mod4.14d <- lda(mpg01 ~ cylinders + displacement + weight, data = train)
mod4.14d.pred <- predict(mod4.14d, test)
mod4.14d.class <- mod4.14d.pred$class
mean(mod4.14d.class != test$mpg01)
```

## (e) Perform QDA on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?

```{r}
mod4.14e <- qda(mpg01 ~ cylinders + displacement + weight, data = train)
mod4.14e.pred <- predict(mod4.14e, test)
mod4.14e.class <- mod4.14e.pred$class
mean(mod4.14e.class != test$mpg01)
```

## (f) Perform logistic regression on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?

```{r}
mod4.14f <- glm(mpg01 ~ cylinders + displacement + weight, family = binomial(), data = train)
mod4.14f.prob <- predict(mod4.14f, test, type = 'response' )
mod4.14f.pred <- ifelse(mod4.14f.prob > 0.5, 1, 0)
mean(mod4.14f.pred != test$mpg01)
```


## (g) Perform naive Bayes on the training data in order to predict mpg01 using the variables that seemed most associated with mpg01 in (b). What is the test error of the model obtained?

```{r}
mod4.14g <- naiveBayes(mpg01 ~ cylinders + displacement + weight, data = train)
mod4.14g.class <- predict(mod4.14g, test)
mean(mod4.14g.class != test$mpg01)
```

## (h) Perform KNN on the training data, with several values of K, in order to predict mpg01. Use only the variables that seemed most associated with mpg01 in (b). What test errors do you obtain? Which value of K seems to perform the best on this data set?

```{r}
train_x <- train %>% select(cylinders, displacement, weight)
test_x <- test %>% select(cylinders, displacement, weight)
train_y <- train$mpg01
test_y <- test$mpg01
for(i in 1:7){
mod4.14h <- knn(train_x, test_x, train_y, k = i)
print(paste('k =', i, ':',mean(mod4.14h != test_y)))
}
```
Based on the results above, k = 6 seems to perform the best in this data set.

## 15. This problem involves writing functions.
## (a) Write a function, Power(), that prints out the result of raising 2 to the 3rd power. In other words, your function should compute 2^3 and print out the results. Hint: Recall that x^a raises x to the power a. Use the print() function to output the result.

```{r}
Power <- function(x){
  print(x*x*x)
}
Power(2)
```

## (b) Create a new function, Power2(), that allows you to pass any two numbers, x and a, and prints out the value of x^a. You can do this by beginning your function with the line. You should be able to call your function by entering, for instance, Power2(3, 8) on the command line. This should output the value of 3^8, namely, 6, 561.
```{r}
Power2 <- function(x, a){
  print(x^a)
}

Power2(3,8)
```

## (c) Using the Power2() function that you just wrote, compute 10^3, 8^17, and 131^3.

```{r}
Power2(10, 3)
Power2(8, 17)
Power2(131, 3)
```

## (d) Now create a new function, Power3(), that actually returns the result x^a as an R object, rather than simply printing it to the screen. That is, if you store the value x^a in an object called result within your function, then you can simply return() this result, using the following line:return(result) The line above should be the last line in your function, before the } symbol.

```{r}
Power3 <- function(x, a){
  result <- x^a
  return(result)
}
```

## (e) Now using the Power3() function, create a plot of f(x) = x^2. The x-axis should display a range of integers from 1 to 10, and the y-axis should display x^2. Label the axes appropriately, and use an appropriate title for the figure. Consider displaying either the x-axis, the y-axis, or both on the log-scale. You can do this by using log = "x", log = "y", or log = "xy" as arguments to the plot() function.

```{r}
plot(x = 1:10, y = Power3(1:10, 2), xlab = 'x', ylab = 'y', main = 'y = x^2')
plot(x = 1:10, y = Power3(1:10, 2), xlab = 'log(x)', ylab = 'y', main = 'y = x^2', log = 'x')
```

## (f) Create a function, PlotPower(), that allows you to create a plot of x against x^a for a fixed a and for a range of values of x. For instance, if you call PlotPower(1:10, 3) then a plot should be created with an x-axis taking on values 1,2,...,10, and a y-axis taking on values 1^3,2^3,...,10^3.

```{r}
PlotPower <- function(x, a){
  return(plot(x, Power3(x,a), xlab = 'x', ylab = 'y'))
}

PlotPower(1:10, 3)
```

## 16. Using the Boston data set, fit classification models in order to predict whether a given census tract has a crime rate above or below the median. Explore logistic regression, LDA, naive Bayes, and KNN models using various subsets of the predictors. Describe your findings. Hint: You will have to create the response variable yourself, using the variables that are contained in the Boston data set.

```{r}
#create a dummy variable 
data(Boston)
boston <- Boston
boston$crim01 <- as.integer(Boston$crim > median(Boston$crim))
```


```{r}
#some visualizations & correlation matrix
pairs(boston[-1])
cor(boston[-1])
```

```{r}
#train test split
set.seed(2022)
sap_size <- floor(0.7 * nrow(boston))
trian_ind <- sample(seq_len(nrow(boston)), size = sap_size)
train <- boston[trian_ind, ]
test <- boston[-train_ind,]
``` 

```{r}
#logistic regression
glm.fit <- glm(crim01 ~ zn*age*dis + indus + nox + rad*tax, family = binomial(), data = train)
summary(glm.fit)
glm.fit.prob <- predict(glm.fit, test, type = 'response')
glm.fit.pred <- ifelse(glm.fit.prob > 0.5, 1, 0)
mean(glm.fit.pred != test$crim01)
```
```{r}
#LDA
lda.fit <- lda(crim01 ~ zn*age*dis + indus + nox + rad + tax, data = train)
lda.fit.pred <- predict(lda.fit, test)
lda.fit.class <- lda.fit.pred$class
mean(lda.fit.class != test$crim01)
```


```{r}
#QDA
qda.fit <- qda(crim01 ~ zn + age + indus + nox + rad + tax, data = train)
qda.fit.pred <- predict(qda.fit, test)
qda.fit.class <- qda.fit.pred$class
mean(qda.fit.class != test$crim01)
```

```{r}
#naive bayes
nb.fit <- naiveBayes(crim01 ~ zn + age + dis + indus + nox + rad + tax, data = train)
nb.fit.pred <- predict(nb.fit, test)
mean(nb.fit.pred != test$crim01)
```

```{r}
#knn
train_x <- train %>% select(zn, age, dis, indus, nox, rad, tax)
test_x <- test %>% select(zn, age, dis, indus, nox, rad, tax)
train_y <- train$crim01
test_y <- test$crim01
for(i in 1:10){
knn.fit <- knn(train_x, test_x, train_y, k = i)
print(paste(i,mean(knn.fit != test_y)))
}
```
It is surprising that KNN(k=1) give us the lowest test error rate. And naive bayes classier returns the highest test error rate. The reason for that might be the predictors are correlated while in the bayes classier, we assume that each predictors are independent. 










